---
title: "Data download"
author: "Anne van der Vliet"
date: "2025-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, include=FALSE}
#load needed packages
library(tidyverse)
library(knitr)
```


# Download dataset
Dataset was downloaded from Dropbox.
```{r warning=FALSE, include=FALSE}
dropbox_url <- "https://www.dropbox.com/scl/fi/3jzhkicf953bsusjtu938/amz_uk_processed_data.csv?rlkey=ikvjt5kz4hy0tbdynrypb786n&st=6xb1y3g6&dl=1"
Amazon_UK <- read_csv(dropbox_url)
```
# Data Description before Data Preparation
For this project, a large dataset from Kaggle (https://www.kaggle.com/) was used, consisting of `r nrow(Amazon_UK)` Amazon products from the UK.

## Remove and rename columns
Various columns were deleted as they will not be used in the analysis. In addition, the needed columns were renamed.
```{r echo=FALSE, warning=FALSE}
#Drop columns
Amazon_UK <- Amazon_UK %>% select(-imgUrl) %>% select(-productURL) %>% select (-reviews) %>% select(-isBestSeller)

#Rename columns
Amazon_UK <- Amazon_UK %>% rename(Product_ID = asin, Product_Name = title, Price = price, SalesLastMonth = boughtInLastMonth, Product_Category = categoryName, StarRating = stars)
```

## Data cleaning
First, prices of 0 were removed as this indicates that the price information was missing. Next, the 1% quantile for each product category, including the extremely low prices, was calculated and these prices were removed for each category.
```{r echo=FALSE, warning=FALSE}
#Prices of 0 removed as this meant no pricing information was available
Amazon_UK <- Amazon_UK %>% filter(Price >= 0.01)

#calculate the 1% quantile for each product category
q01 <- Amazon_UK %>% group_by(Product_Category) %>% summarize(q01=quantile(Price, .01))

#Filter out the extremely low prices
Amazon_UK <- Amazon_UK %>%
  left_join(q01, by = "Product_Category") %>% 
  filter(Price > q01) %>% 
  select(-q01)
```

## Variable operationalization
A price ending variable is created, indicating the price ending of the product (the digits after the whole dollar). In addition, the price strategy variable is created, indicating whether companies used "Just-below pricing", "Round pricing", or another pricing strategy. Finally, price buckets were created, indicating the price level category of the products. 
```{r warning=FALSE, include=FALSE}
#Price ending
Amazon_UK <- Amazon_UK %>%  mutate(HigherPrice = ceiling(Price))
Amazon_UK <- Amazon_UK %>%  mutate(BelowPrice = HigherPrice-Price)
Amazon_UK <- Amazon_UK %>%  mutate(PriceEnding = (1 - BelowPrice))
Amazon_UK$PriceEnding <- as.numeric(Amazon_UK$PriceEnding)

#Just-below vs. Round pricing (+ other)
Amazon_UK <- Amazon_UK %>%  mutate(Price_Strategy = case_when(
    PriceEnding > 0.9999 ~ "round",
    PriceEnding > 0.9899 & PriceEnding < 1.000 ~ "just-below",
    TRUE ~ "other"
  ))

#Price buckets
Amazon_UK = Amazon_UK %>% mutate(Price_Bucket = cut(Price, c(0,1,5,10,20,50,100, 500, 1000,10000000)))

#Top 10 categories
Categories = unique(Amazon_UK$Product_Category)
Categories = Amazon_UK %>% group_by(Product_Category) %>% summarize(Product_count = n()) %>% mutate(ranking = rank(-Product_count)) %>% mutate(top = ranking <= 10)
Categories_filtered = Categories %>% filter(top==T) %>% pull(Product_Category)

#Remove unneeded columns after operationalization
Amazon_UK <- Amazon_UK %>% select(-BelowPrice, -HigherPrice)
```

##Create dataset with only Top 10 product categories
```{r}
Amazon_UK_Top10 <- Amazon_UK %>% filter(Product_Category %in% Categories_filtered)
```


## Data Description after Data Preparation
After data cleaning and variable operationalization, the datasets consist of `r nrow(Amazon_UK)` Amazon products and includes the following variables:

```{r echo=FALSE, warning=FALSE}
#give overview of variables and descriptions
table <- data.frame(`Variable` = c("Product_ID", "Product_Name", "StarRating", "Price", "SalesLastMonth", "Product_Category", "PriceEnding", "Price_Strategy", "Price_Bucket"), 
                    `Description` = c("Product ID from Amazon", 
                                      "Name of the product",
                                      "Average star rating of the product",
                                      "Current price of the product",
                                      "Number of products sold in the last month", 
                                      "Product category",
                                      "Price ending of the product (digits after the whole dollar)", 
                                      "Price strategy used (just-below vs. Round vs. Other",
                                      "Price level category of the product"),
                    stringsAsFactors = FALSE
)

kable(table, format = "markdown")
```

```{r}
#check whether or not the data types are correct 
glimpse(Amazon_UK)
glimpse(Amazon_UK_Top10)
```

```{r warning=FALSE, include=FALSE}
# Create directories to save data
dir.create('../../gen')
dir.create('../../gen/data-preparation')
dir.create('../../gen/data-preparation/data')
```

```{r warning=FALSE, include=FALSE}
#Save dataset as CSV
write.csv(Amazon_UK, file = "../../gen/data-preparation/data/AmazonData.csv")
write.csv(Amazon_UK_Top10, file = "../../gen/data-preparation/data/AmazonDataTop10.csv")
```

