---
title: "Data download"
author: "Anne van der Vliet"
date: "2025-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r}
#load needed packages
library(tidyverse)
library(knitr)
```


## Download dataset
```{r}
dropbox_url <- "https://www.dropbox.com/scl/fi/3jzhkicf953bsusjtu938/amz_uk_processed_data.csv?rlkey=ikvjt5kz4hy0tbdynrypb786n&st=6xb1y3g6&dl=1"
Amazon_UK <- read_csv(dropbox_url)
```
# Data Description before Data Preparation
For this project, a large dataset from Kaggle (https://www.kaggle.com/) was used, consisting of `r nrow(Amazon_UK)` Amazon products from the UK.

### Remove and rename columns
```{r warning=FALSE}
#Drop columns
Amazon_UK <- Amazon_UK %>% select(-imgUrl) %>% select(-productURL) %>% select(-stars) %>% select (-reviews) %>% select(-isBestSeller)

#Rename columns
Amazon_UK <- Amazon_UK %>% rename(Product_ID = asin, Product_Name = title, Price = price, SalesLastMonth = boughtInLastMonth, Product_Category = categoryName)
```

## Data cleaning
```{r warning=FALSE}
#Prices of 0 removed as this meant no pricing information was available
Amazon_UK <- Amazon_UK %>% filter(Price >= 0.01)

#calculate the 1% quantile for each product category
q01 <- Amazon_UK %>% group_by(Product_Category) %>% summarize(q01=quantile(Price, .01))
#Filter out the extremely low prices
Amazon_UK <- Amazon_UK %>%
  left_join(q01, by = "Product_Category") %>% 
  filter(Price > q01) %>% 
  select(-q01)
```

## Variable operationalization
```{r warning=FALSE}
#Price ending
Amazon_UK <- Amazon_UK %>%  mutate(HigherPrice = ceiling(Price))
Amazon_UK <- Amazon_UK %>%  mutate(BelowPrice = HigherPrice-Price)
Amazon_UK <- Amazon_UK %>%  mutate(PriceEnding = (1 - BelowPrice))
Amazon_UK$PriceEnding <- as.numeric(Amazon_UK$PriceEnding)

#Just-below vs. Round pricing (+ other)
Amazon_UK <- Amazon_UK %>%  mutate(Price_Strategy = case_when(
    PriceEnding > 0.9999 ~ "round",
    PriceEnding > 0.9899 & PriceEnding < 1.000 ~ "just-below",
    TRUE ~ "other"
  ))

#Price buckets
Amazon_UK = Amazon_UK %>% mutate(Price_Bucket = cut(Price, c(0,1,5,10,20,50,100, 500, 1000,10000000)))

#Top 10 categories
Categories = unique(Amazon_UK$Product_Category)
Categories = Amazon_UK %>% group_by(Product_Category) %>% summarize(Product_count = n()) %>% mutate(ranking = rank(-Product_count)) %>% mutate(top = ranking <= 10)
Categories_filtered = Categories %>% filter(top==T) %>% pull(Product_Category)

#Remove unneeded columns after operationalization
Amazon_UK <- Amazon_UK %>% select(-BelowPrice, -HigherPrice)
```

# Data Description after Data Preparation
After data cleaning and variable operationalization, the dataset consists of `r nrow(Amazon_UK)` Amazon products and includes the following variables:

```{r echo=FALSE, warning=FALSE}
#give overview of variables and descriptions
table <- data.frame(`Variable` = c("Product_ID", "Product_Name", "Price", "SalesLastMonth", "Product_Category", "PriceEnding", "Price_Strategy", "Price_Bucket"), 
                    `Description` = c("Product ID from Amazon", 
                                      "Name of the product",
                                      "Current price of the product",
                                      "Number of products sold in the last month", 
                                      "Product category",
                                      "Price ending of the product (digits after the whole dollar", 
                                      "Price strategy used (just-below vs. Round vs. Other",
                                      "Price level category of the product"),
                    stringsAsFactors = FALSE
)

kable(table, format = "markdown")
```