---
title: "Main_analysis"
author: "Anne van der Vliet"
date: "2025-03-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, include=FALSE}
# Load needed packages
library(tidyverse)
library(knitr)
library(dplyr)
library(readr)
library(rmarkdown)
library(tinytex)
```

# Dataset
As indicated in the preliminary analysis, as the distribution regarding the use of just-below vs. round is fairly the same for the entire dataset and the top 10 product category dataset, the analysis will be performed on the top 10 product category dataset. 
```{r warning=FALSE, include=FALSE}
# Load dataset based on "Data Download.Rmd"
Amazon_UK_Top10 <- read_csv("../../gen/data-preparation/output/AmazonDataTop10.csv")
```

# Assumptions linear regression
## Normality
```{r}
# Normality check by inspecting the data
ggplot(Amazon_UK_Top10, aes(SalesLastMonth))+ geom_histogram(binwidth = 25)+ xlim(0, 4000) + ylim(0, 2000)
```

# Regression
```{r}
MainLinModel <- lm(SalesLastMonth ~ Price_Strategy + Price_Strategy * Price, data = Amazon_UK_Top10)
summary(MainLinModel)
```
As can be seen above, a significant model was created, F(5, 967648) = 147.7, p < 2.2e-16, R2 = 0.0007627.

## Main effect IV on DV
Regarding the independent variable (Price Strategy), just-below pricing is treated as the baseline. As can be seem from the analysis, all coefficients are significant. Zooming in on the effect of the IV on the DV, the following conclusions can be made:

- When the price ending of a product is other (between .01 and 0.98), compared to a just below price ending (.99), the sales decrease by 3.14 units on average (β=-3.140, t= -13.590, p< .000).
- When the price ending of a product is round (.00), compared to a just below price ending (.99), the sales increase by 6.04 units on average (β=6.040, t= 12.590, p< .000). 

**This indicates that, considering all products on average, it is best to use a round price (.00) to maximize sales.**

## Main effect MOD on DV
Looking at the direct effect of price (MOD) on sales (DV), for every 1 unit increase in price,sales decrease by 0.0054 units (β=-0.005, t= -9.606, p< .000). This effect is significant; however, very small.

**This indicates that, considering all products on average, an increase in price is associated with a small decrease in sales**
  
## Interaction effect IV*MOD on DV
Looking at the interaction term: 
- When the price ending of a product is other (between .01 and 0.98), the effect of price on sales becomes less negative by 0.0027 units in sales per 1 unit increase in price (β=0.0027, t= 4.331, p< .000).
- When the price strategy is round (.00), the effect of price on sales becomes less negative by 0.0018 (β=0.0018, t=2.207, p= .0273).
