---
title: "Preliminary_data_analysis"
author: "Anne van der Vliet"
date: "2025-03-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, include=FALSE}
# Load needed packages
library(tidyverse)
library(knitr)
library(dplyr)
library(readr)
library(rmarkdown)
library(tinytex)
```

# Dataset
As the distribution regarding the use of just-below vs. round is fairly the same for the entire dataset and the top 10 product category dataset, the analysis will be performed on the top 10 product category dataset. 
```{r warning=FALSE, include=FALSE}
# Load dataset based on "Data Download.Rmd"
Amazon_UK_Top10 <- read_csv("../../gen/data-preparation/output/AmazonDataTop10.csv")
```

```{r warning=FALSE, include=FALSE}
# Create directories to save output
dir.create('../../gen/analysis')
dir.create('../../gen/analysis/output')
```

## Descriptive statistics
First we present some descriptive statistics below to get a first feeling of the data. As can be seen from the table, the proportion of products priced as just-below differs per product category. 
```{r echo=FALSE, warning=FALSE}
# Descriptive statistics table
Amazon_UK_Top10 %>% 
  group_by(Product_Category) %>% 
  summarise(AmountofProducts = n(),
            PricedasJustBelow = sum(Price_Strategy == "just-below", na.rm = TRUE),
            PercentageJustBelow = paste0(round((PricedasJustBelow / AmountofProducts) * 100, 0), "%"),
            AveragePrice = paste0("$", (round(mean(Price, na.rm = TRUE), 2))), 
            MaxPrice = paste0("$", (max(Price, na.rm = TRUE))),
            MinPrice = paste0("$", (min(Price, na.rm = TRUE))),
            AverageReviewRating = round(mean(StarRating, na.rm = TRUE), 1))
```

## ANOVA main effect IV-DV
Before the main analysis is performed, we have looked at the direct effect of the IV (price strategy) on the DV (sales) via an ANOVA. The output of the ANOVA is presented below:
```{r echo=FALSE, warning=FALSE}
# ANOVA
anova_IVDV <- aov(SalesLastMonth ~ Price_Strategy, data = Amazon_UK_Top10)
summary(anova_IVDV)

# Post-hoc Test
tukey_resultIVDV <- TukeyHSD(anova_IVDV)
tukey_resultIVDV

# Means table
Amazon_UK_Top10 %>%
  group_by(Price_Strategy) %>%
  summarise(Mean_Sales = mean(SalesLastMonth, na.rm = TRUE))
```
Based on the ANOVA, one can see a significant effect of pricing strategy on sales. Specifically, when looking at the post-hoc test, there is a significant difference between all groups (round, just-below and other). The means table shows that a round price results in the highest sales, after which a just-below price is best. 

This might; however, differ depending on the price of the product. The main analysis will inspect this further. 


## Linear regression price on sales
In addition, before the main analysis, we have looked at the main effect of the MOD (price) on the DV (sales of the product). The output is presented below:
```{r echo=FALSE, warning=FALSE}
# Run regression price level on sales last month
PriceLevelmodel <- lm(SalesLastMonth ~ Price, data = Amazon_UK_Top10)
# View results
summary(PriceLevelmodel)
```
Based on the linear regression, there is a negative relationship visible between price (MOD) and sales (DV). In the main analysis, the interaction between price and pricing strategy will be considered. In addition, this will individually analyzed for every product category.